---
title: " Reproducable Research: Peer Assignment 1 ActivityMonitoring"
author: "Ann Crawford"
date: "March 31, 2017"
output: html_document


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
Read activitydata.csv 

steps    Number of steps takin in a 5-minute interval (missing values are coded as NA) 

date     The date on which the measurement was taken in YYYY-MM-DD format.  This is transformed to date. 

interval Identiﬁer for the 5-minute interval in which measurement was taken


```{r readdata}

rm(list = ls())
##install.packages("stringi")

##install.packages("ggplot2")
##install.packages("dplyr")
##install.packages("magrittr")    ## for %>% pipe operator 
##library(magrittr)
##library(ggplot2)
##library(plyr)

activityraw <-  read.csv("./data/activity.csv")
activity <-activityraw
activity$date <- as.Date(activityraw$date, format="%Y-%m-%d")
```


##What is the mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

2. Calculate and report the mean and median total number of steps taken
per day

```{r stepsperday}

stepst<- aggregate(steps ~ date,  data= activity, sum,  na.rm=TRUE)

meanstepsPerDay <- format(mean(stepst$steps, na.rm = TRUE), scientific = FALSE)

medianstepsPerDay <- median(stepst$steps, na.rm = TRUE)

```
The average steps per day is `r meanstepsPerDay`  .

The median steps per day is `r medianstepsPerDay` .

```{r histstps}
phist <- ggplot(data=stepst, aes(stepst$steps)) + 
  geom_histogram(
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  labs(title="Histogram of Total Steps per Day") +
  labs(x="Total", y="Count") 
phist

```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all days (y-axis)

2. Which 5-minute interval, on average across all the days in the dataset,
contains the maximum number of steps?

```{r dailypattern}

avgint <- aggregate(steps ~ interval,  data= activity, mean,  na.rm=TRUE)


 ggplot (avgint, aes(x=interval,y=steps)) + 
  geom_line() +
  labs(title="Average Steps per Day Over 5 Minute Interval") +
  labs(x="Interval", y="Average Step Count") 
 

maxsteps <- subset(avgint,steps == max(avgint$steps) )

```

The most steps on average are taken in the `r maxsteps[1,1] ` interval. 

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset
(i.e. the total number of rows with NAs)

2. Devise a strategy for filling in all of the missing values in the dataset. The
strategy does not need to be sophisticated. For example, you could use
the mean/median for that day, or the mean for that 5-minute interval, etc.
reference http://publish.illinois.edu/spencer-guerrero/2014/12/11/2-dealing-with-missing-data-in-r-omit-approx-or-spline-part-1/


3. Create a new dataset that is equal to the original dataset but with the
missing data filled in.

4. Make a histogram of the total number of steps taken each day and Calculate
and report the mean and median total number of steps taken per day. Do
these values differ from the estimates from the first part of the assignment?

What is the impact of imputing missing data on the estimates of the total
daily number of steps?
Are there differences in activity patterns

```{r missingvalues}
Totalmissing <- sum(is.na(activityraw$date)) +sum(is.na(activityraw$interval)) + sum(is.na(activityraw$steps))


NoNa$steps = transform(activityraw, steps = ifelse(is.na(steps), mean(steps, na.rm=TRUE), steps))

total2 <- NoNa %>%  group_by(date) %>%
                    summarize(sum)



meanstepsPerDay2 <- format(mean(NoNa$steps), scientific = FALSE)

medianstepsPerDay2 <- median(NoNa$steps)

phist2 <- ggplot(data=stepTotal, aes(stepTotal$steps)) + 
  geom_histogram(
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  labs(title="Histogram of Total Steps per Day Missing Values replaced with average steps") +
  labs(x="Total", y="Count") 
```

##Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday”
and “weekend” indicating whether a given date is a weekday or weekend
day.

2. Make a panel plot containing a time series plot (i.e. type = "l") of the
5-minute interval (x-axis) and the average number of steps taken, averaged
across all weekday days or weekend days (y-axis). 

```{ r weekdays }

activity %>% 
  dayofweek = weekdays(activity$date)




```