---
title: " Reproducable Research: Peer Assignment 1 ActivityMonitoring"
author: "Ann Crawford"
date: "March 31, 2017"
output: html_document


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
Read activitydata.csv 

steps    Number of steps takin in a 5-minute interval (missing values are coded as NA) 

date     The date on which the measurement was taken in YYYY-MM-DD format.  This is transformed to date. 

interval Identiﬁer for the 5-minute interval in which measurement was taken


```{r readdata}

rm(list = ls())

##install.packages("ggplot2")
library(ggplot2)

activity <-  read.csv("./data/activity.csv")
activity$date <- as.Date(activity$date, format="%Y-%m-%d")
```


##What is mean total number of steps taken per day?

```{r stepsperday}

stepst<- aggregate(steps ~ date,  data= activity, sum,  na.rm=TRUE)

meanstepsPerDay <- mean(stepst$steps, na.rm = TRUE)

medianstepsPerDay <- median(stepst$steps, na.rm = TRUE)

```
The average steps per day is `r meanstepsPerDay`  .

The median steps per day is `r medianstepsPerDay` .

```{r histstps}
phist <- ggplot(data=stepst, aes(stepst$steps)) + 
  geom_histogram(
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  labs(title="Histogram of Total Steps per Day") +
  labs(x="Total", y="Count") 
phist

```

## What is the average daily activity pattern?



## Imputing missing values
```{r missingvalues}
sum(is.na(activity$date))
sum(is.na(activity$interval))
sum(is.na(activity$steps))


```

##Are there differences in activity patterns between weekdays and weekends?
